---
title: "HW2"
author: "Abigail Salle"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<br>
<br>
<br>

```{r}
# load necessary packages 
library(tidyverse)
library(readr)
library(tidymodels)
```
<br>
<br>
<br>
```{r}
# read in our data
abalone_data <- read_csv("abalone.csv")
head(abalone_data)
```
<br>
<br>
<br>
Question 1.  
```{r}
# create new variable, age
age <- abalone_data$rings +1.5

# add to set and assess
abalone_data2 <- cbind(abalone_data, age)
head(abalone_data2)
hist(abalone_data2$age, breaks=30, col="pink", xlab="Years")
```
 <br>
  Age appears to be nearly symmetrical and approximately normal. If anything it is slightly right skewed. The maximum peak is around 11 years.  
   <br>
<br>
<br>
<br>
<br>  
 Question 2:  
```{r}
# split data into train and test sets 
set.seed(1114)

abalone_split <- initial_split(abalone_data2, prop = 0.80,
                                strata = age)
abalone_train <- training(abalone_split)
abalone_test <- testing(abalone_split)
```
  <br>
<br>
<br>
<br>
<br>  
 Question 3:
```{r}
# create recipe 
simple_abalone_recipe <- recipe(age ~ 
          type+longest_shell+diameter+height+whole_weight+shucked_weight+viscera_weight, data = abalone_train)
simple_abalone_recipe
```
 <br>
 We won't use rings to predict age because rings are already being considered. We used rings to create the variable "age" and so using it as a predictor variable again would be redundant / not make sense.
<br>
<br>
```{r}
# create better recipe and modify 
abalone_recipe <- recipe(age ~ 
          type+longest_shell+diameter+height+whole_weight+shucked_weight+viscera_weight+shell_weight, data = abalone_train) %>%

step_dummy(all_nominal_predictors()) %>%
step_interact(terms= ~longest_shell:diameter) %>%
step_interact(terms= ~starts_with("type"):shucked_weight) %>%
step_interact(terms= ~shucked_weight:shell_weight) %>%   
step_zv(all_predictors()) %>%
step_normalize(all_numeric(), -all_outcomes()) %>%
prep(verbose = TRUE, log_changes = TRUE)

```
  <br>
<br>
<br>
<br>
<br>  
 Question 4:
```{r}
# linear regression object 
lm_model <- linear_reg() %>% 
  set_engine("lm")
```
 <br>
<br>
<br>
<br>
<br>  
 Question 5:
```{r}
# workflow
lm_wflow <- workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(abalone_recipe)
```
  <br>
<br>
<br>
<br>
<br>  
 Question 6:
```{r}
# try prediction
lm_fit <- fit(lm_wflow, abalone_train)
lm_fit %>% 
  extract_fit_parsnip() %>% 
  tidy()

hypothetical <- data.frame(type = "F", longest_shell = 0.50, diameter = 0.10, height = 0.30, 
                           whole_weight = 4, shucked_weight = 1, viscera_weight = 2, shell_weight = 1)
predict(lm_fit, new_data = hypothetical)
```
<br>
Based on those criteria, the predicted age is 23.103 years.
  <br>
<br>
<br>
<br>
<br>  
 Question 7:
```{r}
# assess model's performance 

# PART 2
# performance of model 
abalone_train_res <- predict(lm_fit, new_data = abalone_train %>% select(-age))
abalone_train_res %>% 
head()

# versus actual.. create tibble of both
abalone_train_res <- bind_cols(abalone_train_res, abalone_train %>% select(age)) 
abalone_train_res %>% 
  head()

# PART 1
# metric set
rmse(abalone_train_res, truth = age, estimate = .pred)
abalone_metrics <- metric_set(rmse, rsq, mae)
abalone_metrics

# PART 3
# apply metric set to tibble
abalone_metrics(abalone_train_res, truth = age, estimate= .pred )

```
R^2 = 0.05570949 which means just over half of this model represents the data.  
 About 56% of the variability is accounted for, which is not much better than a coin flip so this model clearly is not great.
<br>
<br>
<br>





